\begin{thebibliography}{99}

\bibitem{ghamari2022unmanned}
Mohammad Ghamari, M. S. Hossain, and M. Atiquzzaman,
``Unmanned aerial vehicle communications for civil applications: A review,''
\textit{IEEE Access},
vol. 10, pp. 102492--102531, 2022.

\bibitem{rovira2022review}
Arnau Rovira-Sugranes, Josep M. Barcelo-Ordinas, and Jordi Garcia-Vidal,
``A review of AI-enabled routing protocols for UAV networks: Trends, challenges, and future outlook,''
\textit{Ad Hoc Networks},
vol. 130, p. 102790, 2022.

\bibitem{nawaz2021uav}
Haque Nawaz, Muhammad Shahid, and M. A. Khan,
``UAV Communication Networks Issues: A Review,''
\textit{Archives of Computational Methods in Engineering},
vol. 28, no. 3, pp. 1349--1369, 2021.

\bibitem{koubaa2019micro}
Anis Koubaa,
``Micro air vehicle link (MAVLink) in a nutshell: A survey,''
\textit{IEEE Access},
vol. 7, pp. 87658--87680, 2019.

\bibitem{cordill2025comprehensive}
Bryce Cordill, Thomas Nguyen, and David Smith,
``A comprehensive survey of security and privacy in UAV systems,''
\textit{IEEE Access}, 2025.

\bibitem{anwekar2023analysis}
Devashri Anwekar and Sunil Phulre,
``Analysis of Congestion Control Techniques to Improve QoS and Frequent Communication in FANET,''
in \textit{Proceedings of the World Conference on Communication and Computing},
pp. 1--7, 2023.

\bibitem{verma2024adaptive}
Lal Pratap Verma, R. Singh, and P. Kumar,
``Adaptive congestion control in IoT networks: Leveraging one-way delay for enhanced performance,''
\textit{Heliyon},
vol. 10, no. 22, 2024.

\bibitem{mohsan2023unmanned}
Syed Agha Hassnain Mohsan, Muhammad A. Khan, and Shahid Mumtaz,
``Unmanned aerial vehicles (UAVs): Practical aspects, applications, open challenges, security issues, and future trends,''
\textit{Intelligent Service Robotics},
vol. 16, no. 1, pp. 109--137, 2023.

\bibitem{gupta2016surveyuav}
Lalit Gupta, Raj Jain, and Gabor Vaszkun,
``Survey of important issues in UAV communication networks,''
\textit{IEEE Communications Surveys \& Tutorials},
vol. 18, no. 2, pp. 1123--1152, 2016.

\bibitem{bekmezci2013flying}
Ilker Bekmezci, Ozgur Koray Sahingoz, and Şamil Temel,
``Flying Ad-Hoc Networks (FANETs): A survey,''
\textit{Ad Hoc Networks},
vol. 11, no. 3, pp. 1254--1270, 2013.

\bibitem{mozaffari2019tutorial}
Mehdi Mozaffari, Walid Saad, Mehdi Bennis, and Merouane Debbah,
``A tutorial on UAVs for wireless networks: Applications, challenges, and open problems,''
\textit{IEEE Communications Surveys \& Tutorials},
vol. 21, no. 3, pp. 2334--2360, 2019.

\bibitem{perkins1999adhoc}
Charles E. Perkins,
``Ad Hoc Networking,''
Addison-Wesley, 1999.

\bibitem{clausen2003olsr}
Thomas Clausen and Philippe Jacquet,
``Optimized Link State Routing Protocol (OLSR),''
RFC 3626, IETF, 2003.

\bibitem{quaritsch2010networking}
Markus Quaritsch et al.,
``Networked UAVs as aerial sensor network for disaster management applications,''
\textit{Elektrotechnik und Informationstechnik},
vol. 127, no. 3, pp. 56--63, 2010.

\bibitem{zanella2014iot}
Andrea Zanella et al.,
``Internet of Things for smart cities,''
\textit{IEEE Internet of Things Journal},
vol. 1, no. 1, pp. 22--32, 2014.

\bibitem{tanenbaum2011computer}
Andrew S. Tanenbaum and David J. Wetherall,
``Computer Networks,''
Pearson Education, 5th Edition, 2011.

\bibitem{breiman2001randomforest}
Leo Breiman,
``Random Forests,''
\textit{Machine Learning},
vol. 45, no. 1, pp. 5--32, 2001.

\bibitem{khan2020mlnetwork}
Shahid Khan, Imran Khan, and S. Khan,
``Machine learning-based congestion control techniques in communication networks: A survey,''
\textit{Journal of Network and Computer Applications},
vol. 167, p. 102760, 2020.

\bibitem{fotouhi2019survey}
Ahmad Fotouhi et al.,
``Survey on UAV cellular communications: Practical aspects, standardization advancements, regulation, and security challenges,''
\textit{IEEE Communications Surveys \& Tutorials},
vol. 21, no. 4, pp. 3417--3442, 2019.

\bibitem{gupta2021mlrouting}
Harsh Gupta, R. Kumar, and S. Singh,
``Machine learning techniques for routing optimization in UAV networks,''
\textit{Sensors},
vol. 21, no. 14, p. 4735, 2021.

\bibitem{rahman2022uavqos}
Md. Atiqur Rahman et al.,
``QoS-aware routing protocols in UAV networks: A review,''
\textit{Computer Communications},
vol. 185, pp. 145--160, 2022.

\bibitem{li2020congestion}
Yong Li et al.,
``Congestion control mechanisms in wireless networks: A survey,''
\textit{IEEE Access},
vol. 8, pp. 172999--173018, 2020.

\bibitem{li2018prioritizing}
Jun Li et al.,
``Prioritizing-based message scheduling for reliable UAV ad hoc networks,''
\textit{International Journal of Performability Engineering},
vol. 14, no. 9, pp. 2021--2030, 2018.

\bibitem{halder2022dynamic}
Subir Halder et al.,
``Dynamic super round-based distributed task scheduling for UAV networks,''
\textit{IEEE Transactions on Wireless Communications},
vol. 22, no. 2, pp. 1014--1028, 2022.

\bibitem{yang2025intelligent}
Kui Yang et al.,
``Intelligent queue scheduling method for SPMA-based UAV networks,''
\textit{Drones},
vol. 9, no. 8, p. 552, 2025.

\bibitem{nguyen2024acs}
T.-D. Nguyen, T. Pham Van, D. Le, and H. Choo,
``Adaptive Clustering and Scheduling for UAV-Enabled Data Aggregation,''
\textit{Electronics}, vol. 13, no. 16, p. 3322, 2024.

\bibitem{wang2024drllagrangian}
S. Wang and Z. Luo,
``Real-Time Data Collection and Trajectory Scheduling Using a DRL-Lagrangian Framework in Multiple UAVs Collaborative Communication Systems,''
\textit{Remote Sensing}, vol. 16, no. 23, p. 4378, 2024.

\end{thebibliography}